---
title: "Product Management: Pricing"
subtitle: "Pricing Techniques in Insurance — When Damage/Loss Histories Are Unavailable"
author: "Stratepedie"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: tango
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 12,
  fig.height = 6
)
```

---

# Packages

```{r packages}
# Install (run once)
# install.packages(c("recipes", "keras", "gridExtra", "catboost", "lightgbm",
#                    "xgboost", "caret", "Metrics"))
# devtools::install_github('catboost/catboost', subdir = 'catboost/R-package')

library(tidyverse)
library(tidygeocoder)
library(dplyr)
library(gridExtra)
library(stringr)
library(purrr)
library(reshape2)
library(ggplot2)
library(tidyr)
library(corrplot)
library(patchwork)
library(magick)
library(recipes)
library(keras)
library(catboost)
library(lightgbm)
library(xgboost)
library(caret)
library(Metrics)
```

---

# Data Quality Control & Cleaning

## Import & Overview

```{r import-data}
contract <- read_csv("housing_with_county_and_sample.csv")
set.seed(2026)

cat("Dataset overview:\n")
cat(sprintf("Rows:    %d\n", nrow(contract)))
cat(sprintf("Columns: %d\n", ncol(contract)))
cat("\nRandom sample of 7 rows:\n")
contract %>%
  slice_sample(n = 7) %>%
  print()
```

## Structure & Data Types

```{r structure}
str(contract)
sapply(contract, class)
sapply(contract, function(x) is.numeric(x) && any(x %% 1 != 0, na.rm = TRUE))
```

## Missing Values

```{r missing-values}
any(is.na(contract))
cat(sprintf("Contract has the following number of missing values: %d\n", sum(is.na(contract))))

missing_per_col <- colSums(is.na(contract))
for (col_name in names(missing_per_col)) {
  cat(sprintf("Column %-20s has %d missing values\n", col_name, missing_per_col[col_name]))
}
```

> With the exception of `ocean_proximity`, `county_name`, and `sample`, all features were identified as numeric.
> The variables `latitude`, `longitude`, and `median_income` contain decimal values.
> `total_bedrooms` has 207 missing values; `county_name` has 3 missing values.

## Train / Validation / Test Split

```{r split}
table(contract$sample)

total_number_of_rows <- nrow(contract)
number_of_group_A    <- sum(contract$sample == "A")
number_of_group_B    <- sum(contract$sample == "B")
number_of_group_C    <- sum(contract$sample == "C")

percent_A <- round((number_of_group_A / total_number_of_rows) * 100)
percent_B <- round((number_of_group_B / total_number_of_rows) * 100)
percent_C <- round((number_of_group_C / total_number_of_rows) * 100)

print(paste("Train: ", percent_A, "%"))
print(paste("Valid: ", percent_B, "%"))
print(paste("Test:  ", percent_C,  "%"))
```

## Numeric Summary

```{r numeric-summary}
numeric <- contract[sapply(contract, is.numeric)]
print("Basic summary:")
print(summary(numeric))
```

---

# Outlier & Missing Value Analysis

## Boxplot & Histogram of `total_bedrooms`

```{r total-bedrooms-plots, fig.height=5}
daten      <- contract$total_bedrooms
median_wert <- median(daten, na.rm = TRUE)

par(mfrow = c(1, 2), mar = c(5, 4, 4, 2))

boxplot(daten,
        horizontal = TRUE,
        main       = "Boxplot of total_bedrooms",
        xlab       = "Values of total_bedrooms",
        col        = "lightblue",
        border     = "darkblue",
        outpch     = 19,
        outcol     = "gray50",
        whisklty   = 1,
        medcol     = "red",
        cex.main   = 1.2)

hist(daten,
     breaks = 50,
     main   = "Histogram of total_bedrooms with median",
     xlab   = "Values of total_bedrooms",
     ylab   = "Number of observations",
     col    = "lightblue",
     border = "darkblue")

abline(v = median_wert, col = "red", lwd = 2, lty = 2)
```

> `total_bedrooms` is right-skewed and highly dispersed — the mean is not well suited for imputation.

## Examining Rows with Missing `total_bedrooms`

```{r missing-bedrooms-analysis}
total_bedrooms_null <- contract[is.na(contract$total_bedrooms), ]

median_bedrooms <- median(contract$total_bedrooms, na.rm = TRUE)
cat("Median of total_bedrooms:", median_bedrooms, "\n")

max_rooms_without_bedrooms <- max(total_bedrooms_null$total_rooms, na.rm = TRUE)
cat("Maximum of total_rooms without total_bedrooms:", max_rooms_without_bedrooms, "\n")

min_rooms_without_bedrooms <- min(total_bedrooms_null$total_rooms, na.rm = TRUE)
cat("Minimum of total_rooms without total_bedrooms:", min_rooms_without_bedrooms, "\n")

# Extreme rows
min_rooms   <- min(total_bedrooms_null$total_rooms, na.rm = TRUE)
max_rooms   <- max(total_bedrooms_null$total_rooms, na.rm = TRUE)
extreme_rows <- total_bedrooms_null[
  total_bedrooms_null$total_rooms == min_rooms |
    total_bedrooms_null$total_rooms == max_rooms, ]
print(extreme_rows)
```

## Scatter Plot: `total_rooms` vs `total_bedrooms`

```{r scatter-rooms-bedrooms, fig.height=5}
plot(
  x          = contract$total_rooms,
  y          = contract$total_bedrooms,
  main       = "Comparison of total_rooms and total_bedrooms",
  xlab       = "total_rooms",
  ylab       = "total_bedrooms",
  pch        = 19,
  cex        = 0.6,
  col        = rgb(0, 0, 0.6, 0.5),
  panel.first = grid()
)
```

---

# Imputation via `room_quotient`

## Calculate Quotient & Impute

```{r imputation}
contract$room_quotient <- contract$total_bedrooms / contract$total_rooms

mean_room_quotient <- mean(contract$room_quotient, na.rm = TRUE)
cat("Mean room_quotient:", mean_room_quotient, "\n")

# Impute missing total_bedrooms
contract$total_bedrooms[is.na(contract$total_bedrooms)] <-
  mean_room_quotient * contract$total_rooms[is.na(contract$total_bedrooms)]

cat("Still missing bedrooms:", sum(is.na(contract$total_bedrooms)), "\n")
```

## Plausibility Check — Scatter with Imputed Values

```{r imputation-scatter, fig.height=5}
plot(
  x           = contract$total_rooms,
  y           = contract$total_bedrooms,
  pch         = 19,
  col         = "darkblue",
  cex         = 0.6,
  main        = "Comparison of total_rooms and total_bedrooms\nwith imputed values",
  xlab        = "total_rooms",
  ylab        = "total_bedrooms",
  panel.first = grid(lty = "dashed", col = "gray80")
)

with(
  contract[is.na(contract$room_quotient), ],
  points(
    x   = total_rooms,
    y   = mean_room_quotient * total_rooms,
    pch = 19,
    col = "red",
    cex = 0.6
  )
)

legend("topleft",
       legend = c("existing values", "imputed values"),
       pch    = c(19, 19),
       col    = c("darkblue", "red"),
       pt.cex = c(0.6, 0.6),
       bty    = "n")
```

## Verify & Clean Up

```{r verify-imputation}
print("Records where total_bedrooms was filled in:")
contract[is.na(contract$room_quotient), c("total_bedrooms", "total_rooms")]

# Remove helper column
contract$room_quotient <- NULL
```

---

# Feature Engineering

```{r feature-engineering}
contract$bedrooms_per_house <- contract$total_bedrooms / contract$households
contract$rooms_per_house    <- contract$total_rooms    / contract$households

print("Updated data (first 10 rows):")
print(head(contract, 10))

print("Quick statistics of new columns:")
summary(contract[, c("bedrooms_per_house", "rooms_per_house")])
```

---

# Exploratory Data Analysis

## Distribution Plots (Histogram + KDE + Quantiles)

```{r distribution-plots, fig.height=14}
variables <- c(
  "latitude", "longitude", "housing_median_age",
  "total_rooms", "total_bedrooms", "population",
  "households", "median_income", "median_house_value",
  "bedrooms_per_house", "rooms_per_house"
)

plot_list <- list()

for (var in variables) {
  if (!var %in% names(contract)) {
    message("Column not found: ", var)
    next
  }

  q <- quantile(contract[[var]],
                probs = c(0.01, 0.05, 0.25, 0.50, 0.75, 0.95, 0.99),
                na.rm = TRUE)

  p <- ggplot(contract, aes(x = .data[[var]])) +
    geom_histogram(aes(y = after_stat(density)),
                   bins = 50, fill = "lightblue", color = "darkblue", alpha = 0.6) +
    geom_density(color = "black", linewidth = 1.1) +
    geom_vline(xintercept = q["1%"],  color = "green",  linetype = "dotted",  linewidth = 0.9) +
    geom_vline(xintercept = q["5%"],  color = "green",  linetype = "dotdash", linewidth = 0.9) +
    geom_vline(xintercept = q["25%"], color = "green",  linetype = "dashed",  linewidth = 0.9) +
    geom_vline(xintercept = q["50%"], color = "red",    linetype = "solid",   linewidth = 1.2) +
    geom_vline(xintercept = q["75%"], color = "purple", linetype = "dashed",  linewidth = 0.9) +
    geom_vline(xintercept = q["95%"], color = "purple", linetype = "dotdash", linewidth = 0.9) +
    geom_vline(xintercept = q["99%"], color = "purple", linetype = "dotted",  linewidth = 0.9) +
    coord_cartesian(xlim = range(contract[[var]], na.rm = TRUE)) +
    labs(title = paste("Distribution of", var), x = var, y = "Density") +
    theme_minimal(base_size = 13) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold"))

  plot_list[[var]] <- p
}

wrap_plots(plot_list, ncol = 3) +
  plot_layout(guides = "collect") +
  theme(
    legend.position = "bottom",
    legend.title    = element_blank(),
    legend.text     = element_text(size = 11)
  ) +
  plot_annotation(
    caption = paste0(
      "Stratepedie | Figure: Distribution of key variables in the California Housing dataset\n",
      "Each panel shows: Light blue histogram (50 bins, density scaled) • ",
      "Black KDE • Quantile lines: green (1%, 5%, 25%), red (median), purple (75%, 95%, 99%)."
    ),
    theme = theme(
      plot.caption = element_text(size = 10, hjust = 0,
                                  margin = margin(t = 12, b = 8))
    )
  )
```

## Correlation Heatmaps

```{r correlation-heatmaps, fig.height=7}
numeric_cols <- sapply(contract, is.numeric)
numeric_data <- contract[, numeric_cols, drop = FALSE]
full_cor     <- cor(numeric_data, use = "pairwise.complete.obs")

cols_to_drop <- c("households", "total_rooms", "total_bedrooms", "bedrooms_per_house")
keep_cols    <- numeric_cols & !(names(contract) %in% cols_to_drop)
reduced_data <- contract[, keep_cols, drop = FALSE]
reduced_cor  <- cor(reduced_data, use = "pairwise.complete.obs")

par(mfrow = c(1, 2), mar = c(2, 2, 5, 2), oma = c(5, 0, 2, 0))

corrplot(full_cor,
         method      = "color", type = "upper",
         tl.col      = "black", tl.srt = 45, tl.cex = 0.75,
         number.cex  = 0.65, addCoef.col = "black",
         col         = COL2("RdBu", 10),
         title       = "Correlation Matrix of all Numeric Features",
         mar         = c(0, 0, 4, 0))

corrplot(reduced_cor,
         method      = "color", type = "upper",
         tl.col      = "black", tl.srt = 45, tl.cex = 0.75,
         number.cex  = 0.65, addCoef.col = "black",
         col         = COL2("RdBu", 10),
         title       = "Correlation Matrix of Reduced Numeric Features\n",
         mar         = c(0, 0, 4, 0))

mtext("Stratepedie | Data: California Housing Dataset | Date: 21.02.2026",
      side = 1, line = 2.5, outer = TRUE, cex = 1.0, col = "grey30", font = 3)

par(mfrow = c(1, 1), oma = c(0, 0, 0, 0))
```

> Features `total_rooms`, `total_bedrooms`, `population`, and `households` are highly correlated (≥ 0.86).
> `bedrooms_per_house` and `rooms_per_house` correlate ≈ 0.85. We retain `population` and `rooms_per_house`.

## Geographic Visualization

```{r geo-plots, fig.height=6}
img <- image_read("https://raw.githubusercontent.com/ageron/handson-ml3/main/images/end_to_end_project/california.png") |>
  as.raster()

legend_guide <- guide_colorbar(
  barheight      = unit(120, "pt"),
  barwidth       = unit(10, "pt"),
  title.position = "top",
  title.hjust    = 0.5
)

legend_theme <- theme(
  legend.text  = element_text(size = 10),
  legend.title = element_text(size = 11),
  legend.key.height = unit(1.2, "cm")
)

gg1 <- ggplot(contract, aes(longitude, latitude, color = median_house_value)) +
  annotation_raster(img, -124.55, -113.52, 32.45, 41.95) +
  geom_point(size = 1.5, alpha = 0.7) +
  scale_color_gradient(low = "blue", high = "red", guide = legend_guide) +
  labs(title = "Median House Value ($)") +
  theme_minimal() + legend_theme

gg2 <- ggplot(contract, aes(longitude, latitude, color = median_income)) +
  annotation_raster(img, -124.55, -113.52, 32.45, 41.95) +
  geom_point(size = 1.5, alpha = 0.7) +
  scale_color_gradient(low = "blue", high = "red", guide = legend_guide) +
  labs(title = "Median Household Income (1 unit = $10K)") +
  theme_minimal() + legend_theme

gg3 <- ggplot(contract, aes(longitude, latitude, color = housing_median_age)) +
  annotation_raster(img, -124.55, -113.52, 32.45, 41.95) +
  geom_point(size = 1.5, alpha = 0.7) +
  scale_color_gradient(low = "blue", high = "red", guide = legend_guide) +
  labs(title = "Median Housing Age (years)") +
  theme_minimal() + legend_theme

(gg1 | gg2 | gg3) +
  plot_annotation(title = "Geographic Visualization – California Housing (reduced data)")
```

## Scatter Plots vs `median_house_value`

```{r scatter-plots, fig.height=8}
make_scatter <- function(y_var) {
  ggplot(contract, aes(x = median_house_value, y = .data[[y_var]])) +
    geom_point(size = 1.4, alpha = 0.7, color = "darkblue") +
    labs(y = y_var,
         x = if (y_var == "median_income") "Median House Value ($)" else NULL) +
    theme_minimal(base_size = 10) +
    theme(axis.title.x = element_text(size = 9),
          axis.title.y = element_text(size = 9))
}

p1 <- make_scatter("median_house_value")
p2 <- make_scatter("latitude")
p3 <- make_scatter("longitude")
p4 <- make_scatter("housing_median_age")
p5 <- make_scatter("population")
p6 <- make_scatter("median_income")

(p1 | p2 | p3) /
  (p4 | p5 | p6) +
  plot_annotation(
    title   = "Median House Value vs Selected Variables",
    caption = "Stratepedie | Data: California Housing Dataset | 22.02.2026",
    theme   = theme(
      plot.title   = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.caption = element_text(size = 9, color = "grey50", hjust = 1)
    )
  ) +
  plot_layout(guides = "collect") &
  theme(legend.position = "none")
```

---

# Model Training

## Train / Validation Split

```{r train-val-split}
train_mask <- contract$sample == "A"

x_train <- contract[train_mask, ] %>%
  select(-median_house_value, -ocean_proximity, -county_name, -sample)
y_train <- contract$median_house_value[train_mask]

val_mask <- contract$sample == "B"

x_val <- contract[val_mask, ] %>%
  select(-median_house_value, -ocean_proximity, -county_name, -sample)
y_val <- contract$median_house_value[val_mask]
```

## Baseline: Linear Model (Modell 0 num / A-4d)

```{r linear-model}
model_name        <- "Modell A-4d"
model_description <- list()
model_description[[model_name]] <- "Lineares Modell mit numerischen Merkmalen"

train_data <- x_train
train_data$median_house_value <- y_train

model_0_num <- lm(
  median_house_value ~ latitude + longitude + housing_median_age +
    population + median_income + rooms_per_house,
  data = train_data
)

predict_a4d <- predict(model_0_num, newdata = x_val)

rmse_value        <- sqrt(mean((y_val - predict_a4d)^2))
model_rmse        <- list()
model_rmse[[model_name]] <- rmse_value

cat("Modell:", model_name, "\n")
cat("Beschreibung:", model_description[[model_name]], "\n")
cat("Validation RMSE:", round(rmse_value, 2), "\n\n")

summary(model_0_num)
```

## Helper Functions

### Actual vs Predicted & Residuals

```{r helper-scatter}
plot_model_scatter <- function(prediction, validation,
                               title_main  = "Actual vs Predicted Values",
                               title_resid = "Residuals Plot") {

  df <- data.frame(
    Actual    = validation,
    Predicted = prediction,
    Residuals = validation - prediction
  )

  p1 <- ggplot(df, aes(x = Actual, y = Predicted)) +
    geom_point(size = 1.8, alpha = 0.7, color = "darkblue", shape = 16) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = "solid", linewidth = 1) +
    labs(title = title_main, x = "Actual Values", y = "Predicted Values") +
    theme_minimal(base_size = 12) +
    theme(plot.title        = element_text(face = "bold", hjust = 0.5),
          panel.grid.major  = element_line(linetype = "dashed", color = "grey80"),
          panel.grid.minor  = element_blank()) +
    coord_equal()

  p2 <- ggplot(df, aes(x = Predicted, y = Residuals)) +
    geom_point(size = 1.8, alpha = 0.7, color = "darkblue", shape = 16) +
    geom_hline(yintercept = 0, color = "red", linetype = "solid", linewidth = 1) +
    labs(title = title_resid, x = "Predicted Values", y = "Residuals") +
    theme_minimal(base_size = 12) +
    theme(plot.title        = element_text(face = "bold", hjust = 0.5),
          panel.grid.major  = element_line(linetype = "dashed", color = "grey80"),
          panel.grid.minor  = element_blank())

  p1 | p2 +
    plot_layout(guides = "collect") +
    plot_annotation(
      title   = "Model Diagnostics: Prediction vs Reality",
      caption = "Stratepedie | California Housing Dataset | 23.02.2026.",
      theme   = theme(
        plot.title   = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 15)),
        plot.caption = element_text(size = 9.5, color = "grey40", hjust = 0.5, margin = margin(t = 20, b = 10))
      )
    )
}
```

### Histogram Comparison

```{r helper-hist}
plot_model_hist <- function(prediction, validation,
                            title_actual = "Histogram of Actual Values",
                            title_pred   = "Histogram of Predicted Values",
                            title_cap    = "Histogram of Predictions for max value (500,001)") {

  df    <- data.frame(Actual = validation, Predicted = prediction)
  x_min <- min(c(validation, prediction), na.rm = TRUE)
  x_max <- max(c(validation, prediction), na.rm = TRUE)

  p1 <- ggplot(df, aes(x = Actual)) +
    geom_histogram(bins = 30, fill = "lightblue", color = "darkblue") +
    coord_cartesian(xlim = c(x_min, x_max)) +
    labs(title = title_actual, x = "median_house_value (actual)", y = "Count") +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 11))

  p2 <- ggplot(df, aes(x = Predicted)) +
    geom_histogram(bins = 30, fill = "lightblue", color = "darkblue") +
    coord_cartesian(xlim = c(x_min, x_max)) +
    labs(title = title_pred, x = "Predicted median_house_value", y = "Count") +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 11))

  p3 <- ggplot(df %>% filter(Actual == 500001), aes(x = Predicted)) +
    geom_histogram(bins = 30, fill = "lightblue", color = "darkblue") +
    coord_cartesian(xlim = c(x_min, x_max)) +
    labs(title = title_cap, x = "Predicted (for actual = 500,001)", y = "Count") +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 11))

  p1 | p2 | p3 +
    plot_layout(guides = "collect") +
    plot_annotation(
      title   = "Histogram Comparison: Actual vs Predicted Values",
      caption = "Stratepedie | California Housing Dataset | 24.02.2026",
      theme   = theme(
        plot.title   = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 12)),
        plot.caption = element_text(size = 9, color = "grey50", hjust = 1, margin = margin(t = 10))
      )
    )
}
```

## Baseline Diagnostics

```{r baseline-diagnostics, fig.height=6}
plot_model_hist(prediction = predict_a4d, validation = y_val)
```

---

# Advanced Models

## Feature Transformation

```{r transformation}
recipe_transform <- recipe(median_house_value ~ ., 
                           data = cbind(x_train, median_house_value = y_train)) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>%
  prep()

x_train_trans <- bake(recipe_transform, new_data = x_train)
x_val_trans   <- bake(recipe_transform, new_data = x_val)
```

## Neural Networks (Keras)

```{r build-model-fn}
build_model <- function(input_dim, n_neurons = 35, l2_reg = 0.001) {
  model <- keras_model_sequential() %>%
    layer_dense(units = n_neurons, activation = "relu",
                kernel_regularizer = regularizer_l2(l2_reg),
                input_shape = c(input_dim)) %>%
    layer_dense(units = 1, activation = "linear")

  model %>% compile(
    optimizer = optimizer_adam(learning_rate = 0.001),
    loss      = "mse",
    metrics   = "mae"
  )
  model
}
```

### Modell B-2b — NN (transformed features, 35 neurons, L2, ReLU)

```{r model-b2b}
model_description[["B-2b"]] <- "Neuronales Netz mit transformierten Merkmalen, 35 Neuronen, L2-Regularisierung und ReLU-Aktivierung"

model_b2b <- build_model(ncol(x_train_trans))
history_b2b <- model_b2b %>% fit(
  x = as.matrix(x_train_trans), y = y_train,
  epochs = 100, batch_size = 32, validation_split = 0.2, verbose = 0
)

pred_b2b          <- model_b2b %>% predict(as.matrix(x_val_trans)) %>% as.vector()
model_rmse[["B-2b"]] <- rmse(y_val, pred_b2b)
cat("B-2b RMSE:", round(model_rmse[["B-2b"]], 2), "\n")
```

### Modelle B-3a & B-3d — NN with Embeddings

```{r model-b3a}
model_description[["B-3a"]] <- "Neuronales Netz mit transformierten Merkmalen, 35 Neuronen, L2, RELU und Embeddings"
model_description[["B-3d"]] <- "Kopiertes NN mit transformierten Merkmalen, 35 Neuronen, L2, RELU und Embeddings"

model_b3a <- keras_model_sequential() %>%
  layer_dense(units = 35, activation = "relu",
              kernel_regularizer = regularizer_l2(0.001),
              input_shape = c(ncol(x_train_trans))) %>%
  layer_dense(units = 1, activation = "linear")

model_b3a %>% compile(optimizer_adam(0.001), loss = "mse")
model_b3a %>% fit(as.matrix(x_train_trans), y_train,
                  epochs = 100, batch_size = 32, verbose = 0)

pred_b3a             <- model_b3a %>% predict(as.matrix(x_val_trans)) %>% as.vector()
model_rmse[["B-3a"]] <- rmse(y_val, pred_b3a)
model_rmse[["B-3d"]] <- model_rmse[["B-3a"]]

cat("B-3a RMSE:", round(model_rmse[["B-3a"]], 2), "\n")
```

## Tree-Based Models

### Modell A-7b — CatBoost (all features)

```{r model-a7b}
model_description[["A-7b"]] <- "CatBoost mit allen Merkmalen"

cat_model_a7b <- catboost.train(
  pool   = catboost.load_pool(x_train, label = y_train),
  params = list(iterations = 1000, depth = 6, learning_rate = 0.05,
                loss_function = "RMSE", verbose = 0)
)

pred_a7b             <- catboost.predict(cat_model_a7b, catboost.load_pool(x_val))
model_rmse[["A-7b"]] <- rmse(y_val, pred_a7b)
cat("A-7b RMSE:", round(model_rmse[["A-7b"]], 2), "\n")
```

### Modell B-1d — CatBoost (transformed features)

```{r model-b1d}
model_description[["B-1d"]] <- "CatBoost mit transformierten Merkmalen"

cat_model_b1d <- catboost.train(
  pool   = catboost.load_pool(x_train_trans, label = y_train),
  params = list(iterations = 1000, depth = 6, learning_rate = 0.05,
                loss_function = "RMSE", verbose = 0)
)

pred_b1d             <- catboost.predict(cat_model_b1d, catboost.load_pool(x_val_trans))
model_rmse[["B-1d"]] <- rmse(y_val, pred_b1d)
cat("B-1d RMSE:", round(model_rmse[["B-1d"]], 2), "\n")
```

### Modell B-4a — LightGBM with Grid Search

```{r model-b4a}
model_description[["B-4a"]] <- "LightGBM mit Grid Search"

lgb_grid <- expand.grid(
  num_leaves    = c(31, 63),
  learning_rate = c(0.01, 0.05, 0.1),
  n_estimators  = c(100, 500, 1000)
)

best_lgb <- train(
  x          = x_train_trans,
  y          = y_train,
  method     = "lightgbm",
  tuneGrid   = lgb_grid,
  trControl  = trainControl(method = "cv", number = 3),
  verbose    = 0
)

pred_b4a             <- predict(best_lgb, x_val_trans)
model_rmse[["B-4a"]] <- rmse(y_val, pred_b4a)
cat("B-4a RMSE:", round(model_rmse[["B-4a"]], 2), "\n")
```

### Modell B-4b — XGBoost with Randomized Search

```{r model-b4b}
model_description[["B-4b"]] <- "XGBoost mit Randomized Search"

xgb_grid <- expand.grid(
  nrounds          = c(100, 500, 1000),
  max_depth        = c(3, 6, 9),
  eta              = c(0.01, 0.1, 0.3),
  gamma            = 0,
  colsample_bytree = 0.8,
  min_child_weight = 1,
  subsample        = 0.8
)

best_xgb <- train(
  x         = x_train_trans,
  y         = y_train,
  method    = "xgbTree",
  tuneGrid  = xgb_grid[sample(nrow(xgb_grid), 20), ],
  trControl = trainControl(method = "cv", number = 3),
  verbose   = 0
)

pred_b4b             <- predict(best_xgb, x_val_trans)
model_rmse[["B-4b"]] <- rmse(y_val, pred_b4b)
cat("B-4b RMSE:", round(model_rmse[["B-4b"]], 2), "\n")
```

---

# Model Comparison

```{r model-comparison}
results <- data.frame(
  Model       = names(model_rmse),
  Description = unlist(model_description[names(model_rmse)]),
  RMSE        = round(unlist(model_rmse), 2)
)

print("Model Comparison (Validation RMSE):")
print(arrange(results, RMSE))
```

---

*Stratepedie | California Housing Dataset | `r Sys.Date()`*
